---
title: "HW 8"
author: "Lindsay Payne - UTEID: Lnp832"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document:
    toc: true
---

```{r global_options, echo=FALSE}
knitr::opts_chunk$set(fig.height=3, fig.width=4, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60))
```

----------------------
#### GitHub Repository: https://github.com/lindsaypayne017/SDS315HW8.git

### Problem 1: Regression Warm Up
```{r, echo = FALSE, message = FALSE}
# Load in necessary packages
library(tidyverse)
library(mosaic)
```

```{r, echo = FALSE}
# Read in the creatinine data set
creatinine <- read.csv("creatinine.csv")
```

#### Part A
```{r, echo = FALSE, message = FALSE, results = 'hide'}
# Create a linear model for the data
model_creat = lm(creatclear ~ age, data=creatinine)

# Find the coefficients (intercept and slope)
coef(model_creat)
```

```{r, echo = FALSE, message = FALSE, results = 'hide'}
# Calculate the expected creatinine clearance rate of a 55 year old
148 - 0.6 * (55)
```

We would expect a 115 mL/minute creatinine clearance rate for a 55-year-old. To determine this, I fit a regression model using the linear model (lm) function in R on the creatinine data set with "creatclear" as the target variable and "age" as the feature variable before printing the weights/coefficients of the model to the console, giving me the baseline (148) and weight on age (-0.6) needed to create the fitted model ŷ = 148 - 0.6x. I then plugged in 55 as x, or age, into the fitted equation E(CCR | Age = 55) = 148 − 0.6 * 55 and used R to calculate a predicted y of 115 mL/minute.

#### Part B
A one-year change in age is associated with a -0.6 mL/minute change (or -0.6 mL/minute per year) in creatinine clearance rate, on average. To determine this, I fit a regression model using the linear model (lm) function in R on the creatinine data set with "creatclear" as the target variable and "age" as the feature variable. I then printed the weights/coefficients of the fitted model to the console which gave me the baseline of 148 and weight on age, representing the average change in clearance rate for each year increase in age, of -0.6.

#### Part C
```{r, echo = FALSE, results = 'hide'}
# Calculate the creatinine clearance rates of each person using R
148 - 0.6 *(40)
148 - 0.6 *(60)
```

The 40-year-old's creatinine clearance rate is healthier/higher for their age. To determine this, I compared the two patients' residuals or differences between their actual outcomes and expected outcomes. Using the fitted regression model, I calculated the 40-year-old's predicted creatinine clearance rate as ŷ = 148 - 0.6 * 40 = 124. The 40-year-old's actual rate of 135 is above average for their age by 11 mL/minute which I found using the equation Residual = Actual − Predicted = 135 - 124. Meanwhile, the 60-year-old's predicted creatine clearance rate was ŷ = 148 - 0.6 * 60 = 112. Their actual rate of 112 mL/minute is the same as the average clearance rate for their age, as seen by 112 - 112 = 0. Therefore, the 40-year-old had the healthier creatinine clearance rate for their age.

### Problem 2: Model Disease Growth
```{r, echo = FALSE}
# Read in the covid data set
covid <- read.csv("covid.csv")
```

#### Part 1
```{r, echo = FALSE, message = FALSE, results = 'hide'}
# Filter by Italy
covid_italy = covid %>%
  filter(country == 'Italy')

# Find a model of log(y) and x
lm_covid_italy = lm(log(deaths) ~ days_since_first_death, data=covid_italy)
coef(lm_covid_italy)

# Find the doubling time using "the rule of 70"
70/18.3
```

```{r, echo = FALSE, message = FALSE, results = 'hide'}
# Bootstrap the Italy model
boot_covid_italy = do (10000)*lm(log(deaths) ~ days_since_first_death, data=resample(covid_italy))

# Find the 95% confidence interval of the bootrapped Italy model
confint(boot_covid_italy, level = 0.95)

# Second way to do it, but specifically boostrapping the Italy coefficients
#boot_covid_lmitaly = do (10000)*coef(lm(log(deaths) ~ days_since_first_death, #data=resample(covid_italy)))
# 95% confidence interval of Italy coefficients
#confint(boot_covid_lmitaly, level = 0.95)
```

```{r, echo = FALSE, message = FALSE, results = 'hide'}
# Extract the bootstrapped slope estimates or growth rates
boot_growth_italy = boot_covid_italy %>%
  select(days_since_first_death)

# Make a doubling time variable using the rule of 70 on the growth rates
boot_doubling_italy = boot_growth_italy %>%
  mutate(doubling_time = 70 / (days_since_first_death * 100))

# Find the 95% bootstrapped confidence interval for doubling time
quantile(boot_doubling_italy$doubling_time, probs = c(0.025, 0.975))
```

The estimated growth rate of reported daily deaths in Italy is about 18.3% or 0.183 with a 95% bootstrapped confidence interval of (0.159, 0.208). Since 0 is not included in this interval, the growth rate is statistically significant and unlikely to have occurred by chance alone. Additionally, Italy's doubling time is about 3.8 days which I calculated using the doubling time equation DT ≈ 70/r ≈ 70/18.3 ≈ 3.8. The 95% bootstrapped confidence interval of this estimate is (3.4, 4.4). Since 0 is not included, the doubling time is also statistically significant and unlikely the result of chance only.

#### Part 2
```{r, echo = FALSE, message = FALSE, results = 'hide'}
# Filter by Spain
covid_spain = covid %>%
  filter(country == 'Spain')

# Find a model of log(y) and x
lm_covid_spain = lm(log(deaths) ~ days_since_first_death, data=covid_spain)
coef(lm_covid_spain)

# Find the doubling time using "the rule of 70"
70/27.6
```

```{r, echo = FALSE, message = FALSE, results = 'hide'}
# Bootstrap the Spain model
boot_covid_spain = do (10000)*lm(log(deaths) ~ days_since_first_death, data=resample(covid_spain))

# Find the 95% confidence interval of the bootrapped Spain model
confint(boot_covid_spain, level = 0.95)

# Second way to do it, but specifically boostrapping the Spain coefficients
#boot_covid_lmspain = do (10000)*coef(lm(log(deaths) ~ days_since_first_death, data=resample(covid_spain)))
# 95% confidence interval of Italy coefficients
#confint(boot_covid_lmspain, level = 0.95)
```

```{r, echo = FALSE, message = FALSE, results = 'hide'}
# Extract the bootstrapped slope estimates or growth rates
boot_growth_spain = boot_covid_spain %>%
  select(days_since_first_death)

# Make a doubling time variable using the rule of 70 on the growth rates
boot_doubling_spain = boot_growth_spain %>%
  mutate(doubling_time = 70 / (days_since_first_death * 100))

# Find the 95% bootstrapped confidence interval for doubling time
quantile(boot_doubling_spain$doubling_time, probs = c(0.025, 0.975))
```

The estimated growth rate for Spain is 27.6% or 0.276 with a 95% bootstrapped confidence interval of (0.235, 0.318). Since 0 is not included in this interval, the growth rate is statistically significant and unlikely to have occurred by chance alone. Furthermore, Spain's doubling time is about 2.5 days with a 95% bootstrapped confidence interval of (2.2, 3.0). As seen by the absence of 0 in this interval, the doubling time is also statistically significant and unlikely the result of chance only.

#### Part 3
```{r, echo = FALSE, results = 'hide', fig.width = 6, fig.height = 4}
ggplot(covid) + 
  geom_line(aes(x = days_since_first_death, y = deaths, color = country)) +
  scale_color_brewer(type="qual") +
  labs(
    title = "Reported Daily COVID-19 Deaths Over Time",
    x = "Days Since First Reported Death",
    y = "Daily Reported Deaths",
    color = "Country"
  ) +
  theme_minimal()
```

### Problem 3: Price Elasticity of Demand
```{r, echo = FALSE}
# Read in the milk data set
milk <- read.csv("milk.csv")
```

```{r, echo = FALSE, results = 'hide'}
# 
lm_milk = lm(log(sales) ~ log(price), data=milk)
coef(lm_milk)

# Bootstrap the milk model
boot_lm_milk = do (10000)*lm(log(sales) ~ log(price), data=resample(milk))

# Find the 95% confidence interval of the bootrapped milk model
confint(boot_lm_milk, level = 0.95)
```

In light of the data, the estimated price elasticity of demand for milk is -1.62. This means that across these participants, a 1% change in milk price produces a -1.62% change in sales on average, or that consumers want to buy about 1.62% less of it on average. Additionally, the 95% bootstrapped confidence interval of this estimate is (-1.77, -1.46). Because 0 is not present in the interval, the elasticity is statistically significant and unlikely to have occurred by chance alone.

To estimate the elasticity, I first diagnosed a power law by looking for a linear relationship between the price and sales variables on the log-log scale. I then fit a power law by running a regression on the log-log scale, meaning with the log of both price and sales, using the linear model (lm) function in R. Next, I used the coefficient function (coef) to extract the intercept and log(price), which represents elasticity, estimates. These values could also be used to fit a power law by hand, as seen by sales = 110 ⋅ price^(-1.62). Lastly, I found the 95% bootstrapped confidence interval of the elasticity estimate, again using R.
